### Install packages

```{r setpup}
#chunk runs everytime we need it 
#install.packages("dplyr")
library(dplyr)
library(ggplot2)
```

WE will be practicing with `dplyr` functions using the `star-wars` data set

```{r}
starwars
#previewing the top data from the dataset
head(starwars)
#questions about the data-set 
?starwars
```

### Subset rows with `filter`

`filter()` allows us to subset a data-set based on the value

```{r}
# Filter for rows, with human star-wars characters
filter(starwars, species == "Human")

# Save to a new r-object 
human <- filter(starwars, species == "Human")
human 

# FIlter Multiple colloumns at once 
filter(starwars, species == "Human", sex =="male")
```

We can utilize different comparison operators for `filter()`

-   `!` is "not"

-   `==` is "equal"

-   `&` is "and"

-   `|` is "or"

-   `<` is less then

-   `>`greater than

-   `is.na()` returns TRUE if values is NA

```{r}
## return chracters with a missing mass 
filter(starwars, is.na(mass))
```

```{r}
## How to get characters that are NOT human
filter(starwars, species != "Human")
```

variable_name `%in%` vector_of_values returns TRUE if the value of the variable_name is equal ro one of the values in vector_of_values

```{r}
#examples 
skin_color <- "grey"
#returns fales beacues "grey" is not in the vecotr of colours
skin_color %in% c("yellow", "orange", "red")

#How wwould we filter for chracters with brown or blong hair?

filter(starwars, hair_color %in% c("blond", "brown"))
```

#### Utilizing `grepl()`

grepl("characters_of_interest", variable_name) returns TRUE if the value of variable_name contains the characters_of_interest.

```{r}
filter(starwars, grepl("blond", hair_color))

#How would we get all characters with blue in their eyes 
filter(starwars, grepl("blue", eye_color))
```

#### Utilizing comparisons

You can use `>`, `<`, `<=`, `>=` to compare numeric variables or categorical variables

-   nominal categorical variables (character vectors) are ranked alphabetically

-   ordinal categorical variables (factors) are ranked based on the built-in-order

```{r}
## starwars characters with a name starting with the letters befors "D" 
filter(starwars, name < "D")

str(diamonds)
filter (diamonds, cut > "Good") # ranking is based upon the levels of the orderd factor cut 
```

### `Between()`

-   between(x, left, right) is a shorthand for x\>= left & x\<= right

```{r}
between(5,1,7)# true becaues 5 is less than or equal to 7 or greater or equal to 1 
filter(starwars, between(mass, 50, 100))
```

### Practice

```{r}
# Missing hair colour 
filter(starwars, is.na(hair_color))
```

```{r}
# Shorter than 100cm and are not droirds 
filter(starwars, height < 100 & species != "Droid")
```

```{r}
# Born between the years 100 & 200 
filter(starwars, between(birth_year, 100, 200))
```

```{r}
# My choice : characters that where in the orignal triology, but are driods
filter(starwars, gender != "masculine" & species == "Droid")
```

## Order rows with `arange()`

`arange()`reorder rows based on the values of coloum(s)

```{r}
## order by character names 
arrange(starwars, name)

# order of charatcer values 

# order by species and birth year 
arrange(starwars, species, birth_year)

# use desc() to get the  to ge the birth years in descending order 
arrange(starwars, species, desc(birth_year))
```

#### Practice

```{r}
# Arrange tallest to shortest 
arrange(starwars, desc(height))
```

```{r}
# shortest to tallest 
arrange(starwars, height, species)
```

```{r}
arrange(starwars, hair_color, eye_color)
```

### Subset columns with `select()`

```{r}
select(starwars, hair_color, skin_color, eye_color)
character_color <- select(starwars, hair_color, skin_color, eye_color)
character_color
```

we can also select multiple columns in a row

```{r}
select(starwars, name:eye_color)
```

we can also remove columns by name using a `-`

```{r}
starwars
select(starwars, -c(films:starships)) 
```

### Helpful functions

`everything()` is a shortcut for "everything e;se"

```{r}
select(starwars, species, everything()) # put species first, then all other colunms
## alphabetizes columns
starwars_names <- sort(names(starwars))
starwars_names
select(starwars, starwars_names)
```

`starts_with("abc")`matches columns namese that start with "abc"

`ends_with("abc")` matches columns names that end with "abc"

`contains("abc")` matches column names that contain "abc"

```{r}
select(starwars, starts_with("h"))
select(starwars, contains("or"))
```

#### Practice :3

1.  Select each character home-world & species, what question might this help answer

    ```{r}
    select(starwars, species, homeworld, everything())
    ```

-   Where do Humans in the star wars universe most reside?

2.  How many ways to select `hair_color`& `skin_color`

```{{r}}
# way 1 
```

### Add new variables with `mutate()`

```{r}
starwars_small <- select(starwars, height, mass, species)
starwars_small

mutate(starwars_small, height_m = height/100) # mutating the height into meters 
```

we can also create multiple variables at once and we can refer to variables we just created

```{r}
mutate(starwars_small,
       height_m = height/100,
       bmi = mass/height_m^2)
```

Summary statistics give one value for the entire column `sum()`, `mean()`

```{r}
mutate(starwars_small,
       sum = sum(mass, na.rm =  T))
mutate(starwars_small,
       row_addition = height + mass,
       summarry_sum = sum(height, mass, na.rm = T))
```

Group by variable using the `.by` parameter

```{r}
mutate(starwars_small,
       species_mass = mean(mass, na.rm = T),
       .by = species)
```

change types of a variables

```{r}
mutate(starwars_small, 
       mass = as.character(mass))
```

#### Practice

```{r}
# change mass to pounds 
mutate(starwars_small,
       pounds = mass*2.2)
```

### Summarize data with `sumarize()`

the only coloumns kept are the columns you group by and the coloumns you create

```{r}
summarize(starwars, 
          mean_mass = mean(mass, na.rm = T))
# Group by species 
summarize(starwars,
          mean_mass = mean(mass, na.rm = T),
          mean_pounds = mean_mass * 2.2,
          .by = species)
```

`n()` is a helpful function returns the numbers of rows in a group

```{r}
summarize(starwars, 
          mean_mass = mean(mass, na.rm = T),
          cout = n(), 
          .by = species)

## group by multiple groups 
summarize(starwars, 
          mean_mass = mean(mass, na.rm = T),
          cout = n(), 
          .by = c(species, sex))

```

#### Renaming data 

rename(data, new_name = old_name)

```{r}
new_starwars <- rename(starwars, birthYear = birth_year)
```
