### load packages + libraries

```{r}
 # load required packages
 library(tidyr)
 library(janitor)
 library(dplyr)

# pokemon library 
Pokemon = read.csv("/Users/tamiaabbott/Desktop/programing/camps/DSRP/DSRP_2023/Data/Pokemon.csv")

View(Pokemon)
```

## Part 1 - Test for a significant difference between 2 groups 

```{r}
highest_tot = slice_max(Pokemon, n=36, Pokemon$Total, with_ties = F )
highest_tot
```

#### Comparing legendary & non legendary Pokemon based on which ones have the higher total stats

Hypothesis : Legendary Pokemon on average are going to have a higher total stats

Null Hypothesis : Legendary Pokemon and non-legendary Pokemon share the same mean total stats

Alt hypothesis: Legendary Pokemon and non-legendary Pokemon have differing mean total stats

```{r}
# sort based on legendary status 
pkLegndary = Pokemon |> filter(Legendary == "True")
pknonLegndary = Pokemon |> filter(Legendary == "False")

# testing time! 

t.test(pkLegndary$Total, pknonLegndary$Total, paired = F, alternative = "two.sided")
```

**Results**

-   P-value = 2.2 \* 10^-16^
    -   There is a very unlikely that this result was by chance.
    -   
