### load packages + libraries

```{r}
 # load required packages
 library(tidyr)
 library(janitor)
 library(dplyr)

# pokemon library 
Pokemon = read.csv("/Users/tamiaabbott/Desktop/programing/camps/DSRP/DSRP_2023/Data/Pokemon.csv")

View(Pokemon)
```

## Part 1 - Test for a significant difference between 2 groups 

```{r}
highest_tot = slice_max(Pokemon, n=36, Pokemon$Total, with_ties = F )
highest_tot
```

#### Comparing legendary & non legendary Pokemon based on which ones have the higher total stats

Hypothesis : Legendary Pokemon on average are going to have a higher total stats

Null Hypothesis : Legendary Pokemon and non-legendary Pokemon share the same mean total stats

Alt hypothesis: Legendary Pokemon and non-legendary Pokemon have differing mean total stats

```{r}
# sort based on legendary status 
pkLegndary = Pokemon |> filter(Legendary == "True")
pknonLegndary = Pokemon |> filter(Legendary == "False")

# testing time! 

t.test(pkLegndary$Total, pknonLegndary$Total, paired = F, alternative = "two.sided")
```

**Results**

-   P-value = 2.2 \* 10^-16^
    -   There is a very unlikely that this result was by chance.

## Part 2 - **Test for a significant difference between 3+ groups**

#### Looking for a difference for the mean total stats for different Pokemon types

Null Hypothesis : There is no difference in mean total stats for Pokemon from different types

Alt hypothesis: There is a difference in mean total stats for Pokemon from differing types

```{r Part_2}
pk_tytot_results = aov(data = Pokemon, Pokemon$Total ~ Pokemon$Type.1)

## looking at the results 
summary(pk_tytot_results)

## which types are different from one another 
results = TukeyHSD(pk_tytot_results)
pk_res = results$`Pokemon$Type.1`
l = as.data.frame(pk_res)
```

```{r}
slice_max(l, diff, n =1)
```

p-value = 2.08e^-09^

meaning: That there is a apparent difference

## Part 3 - **Test for a significant association between categorical variables**

#### Looking for a relationship between a Pokemon type and if they are a legendary 

Null Hypothesis : There is no relationship between the Pokemon type and whether or not they're a legendary

Alt hypothesis: There is an relationship between the Pokemon type and whether or not they're a legendary.

```{r}
# Cleaning & Preping data 
Pk_clean = Pokemon |> 
  filter(Legendary == "False")

## checking the plots 
ggplot(Pokemon, aes(x = Type.1,  y = Legendary)) + 
  geom_count() +
  theme_minimal()
```

```{r}
# making a contingency table 
t_pk = table(Pokemon$Type.1, Pokemon$Legendary)



# chi - sqaured test 
chi = chisq.test(t_pk)
chi$p.value
chi$residuals

# plotting res 
corrplot(chi$residuals, is.corr = F)
```
