### install + load packages

```{r setup}
 #install.packages("janitor")
 #install.packages("tidyr")
 
 # load required packages
 library(tidyr)
 library(janitor)
 library(dplyr)

```

### Janitor

```{r}
clean_names (clean_names(starwars, case = "small_camel")) 
new_starwars = clean_names(starwars, case = "screaming_snake")

```

### Challenge problem! 

```{r}
Starwars_women = select(arrange(filter(starwars, sex == "female"), birth_year), name, species)
Starwars_women
```

### Pipe \|\>

allows to nestle functions within one another

`function(data, parameters)` becomes `data |> function(parameters)`

```{r}
# using pipes 
sWW = starwars |> 
  filter(sex =="female") |>
  arrange(birth_year) |> 
  select(name, species)

sWW
```

### Slice\_ 

takes rows based on their potions, There are 5 handy functions that allow you to extract specific rows within each group:

-   `slice_head(data, n = 1)` takes the first row from each group

-   `slice_tail(data, n = 1)` takes the last rows in each group

-   `slice_min(data, x, n = 1)` takes the row with the smallest value of column x

-   `slice_max(data, x, n = 1)` takes the row with the largest value of column x

-    `slice_sample(data, n = 1)` takes on random row.

You can vary `n` to select more than one row, or instead of `n=`, you can use `prop = 0.1` to select (e.g) 10% of the rows in each group

```{r}
# 10 tallest starwars characters /w slice 
tall_star_wars = slice_max(starwars, height, n=10)
tall_star_wars

# top 2 of each species 
slice_max(starwars, height, n=10, by = species, with_ties = F) # with ties elimnates ties
```

## Tidy

### pivot_longer 

```{r}
table4a

#Changes columns  + names
tidy_table4a = pivot_longer(table4a, 
             cols = c(`1999`, `2000`),
             names_to = "year",
             values_to = "cases")
```

```{r}
# table b into tidy format 
table4b

tidy_table4b = pivot_longer(table4b,
             cols = c(`1999`, `2000`),
             names_to = "year",
             values_to = "population")
```

### pivot_wider 

```{r}
table2

pivot_wider(table2,
            names_from = type,
            values_from = count)
```

### separate

to split values into 2 different columns

```{r}
table3

separate(table3, 
         rate,
         into = c("cases", "population"),
         sep = "/") #specify when to separate
```

### unite 

joining 2 different columns into 1

```{r}
table5

unite(table5,
      "year",
      col = c("century", "year"),
      sep = "")
# fixing table 5 
tidy_table5 <- table5 |>
  unite("year", col())
```

### bind rows

```{r}
new_data <- data.frame(country = "USA", year = "1999", cases = "1042", population = "200000000")
```
